%! TEX root = main.tex

%/========== Acknowledgements ==========/%

\begin{acknowledgements}
    Manfredi, you are \textit{without a doubt} the greatest supervisor, the
    best mentor, and one of the most wonderful people in the world.
    Knowing you, you'll likely respond to that by saying 
    ``thank you Adan, you're too kind", and then proceed to laugh uncomfortably before
    changing the topic.
    But it's true, and I don't know how else to tell you how lucky and grateful I am to
    have you in my life.

    These two years of research together were the most fun I've ever
    had; they were also fraught with death, sorrow, and the most
    devastating heartbreak I've ever experienced.
    Through it all you've been nothing but kind, honest, generous,
    compassionate, understanding, and so incredibly humble.
    You work tirelessly in the background to ensure your students
    succeed, yet ask for so little in return.
    My name is on the title page of this thesis, but if not for the required
    academic formality I would change the authorship to read
    ``Adan Moran-MacDonald, who couldn't have done any of this without the love
    and support of his supervisor and dear friend, Manfredi Maggiore."

    Manfredi: you're my hero. 

    \bigskip % big line break to separate

    My gratitude also goes to Lacra Pavel, Raymond Kwong, and Stark Draper for
    being on my committee. I hope you enjoyed reading this thesis. Thank you for
    your hard work.

    Dear friends in the Systems Control Group,
    you remind me what it means to have a zest for life. 
    Every moment with you is wild, crazy, never-ending fun. Thank you
    for the memes, deep chats, pub games, and bean juice; the distractions
    always came when I needed them most.

    Oh family of mine, your unconditional support has shaped every word I've
    written in this thesis, and you deserve more appreciation than I can put
    into words. Thank you for driving me to be myself. I love you all.

    Finally, to Mallory, Leela, and Fry: our life together may have ended, but
    you filled my heart and helped me become who I am today. 
    Thank you for the happiest years of my life.

\end{acknowledgements}

%/========== /Acknowledgements ==========/%
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :
